Learning and comparing functional connectomes across subjects

### 2012-12-31 paper.tex
Start of clean up of predictive modelling section

\section{Predictive Modeling}
craddock 2009, Dosenbach 2010, lord 2011, deshpande,

1. importance of interpretation!
2. the need for feature extraction
3. the need for feature selection
4. setting parameters.
\section{Predictive Modeling}
craddock 2009, Dosenbach 2010, lord 2011, deshpande,

@@

@@

Nearly every analytical method can be appplied in a predictive modelling framework, but
multivariate classification and regression have been the most commonly used
(referred to as MVPA; multivariate prediction analysis).  Multivariate methods
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisions since
the significance of an entire pattern is evaluated using a single statistical
test. Predictive modeling is typically performed in a supervised learning
framework where a target is available for each observation of the features,
although unsupervised and semi-supervised methods can also be used.

@@

Model training can be further broken down into feature extraction, model
selection and feature selection and testing can be performed for model
optimziation and

@@

Determining the subset of features most relevant to the model (feature
selection) is essential for interpreting the models, and is the subject of
ongoing research. In cases were model weights can be directly mapped to their
corresponding feature, there is typically no statistical framework for
thresholding these weights, and meaningless features are not always set to zero
(this is particularly the case for L2 regularized algorithms such as SVM and
ridge regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest using a
univariate statistical criterion such as t-test, Pearson's correlation, or
mutual information.  A drawback of this approach is that since the filters are
typically insensitive to multivaratiate information, they may exclude features
that would be important to the model.  Another approach is to treat feature
selection as a model selection problem in which case the features are
iteratively pruned and a subset is chosen based on prediction performance. A
third approach is to use sparse modeling methods such as L1 SVM or LASSO that
force meaningless weights to be zero.  Although ideal, these methods may have
limits on the number of non-zero weights used in the model or the exclusion of
highly correlated features from the model.  Contrary to previous reports [Guyon
2002, Craddock 2009] no matter which feature selection strategy is employed, it
must be performed withen cross-validation to avoid biasing estimates of
prediction accuracy [later Guyon].

The interpretation of classifier weights for selected features, when available,
is complicated and requires insight into the mechanisms underlying the modeling
approach. If features are appropriately standardized prior to training, the
magnitude of weights can be interpreted to reflect the relative importance of
the feature to the model, but interpreting how the feature differs between
classes or relates to a continuous variable is more complicated, particularly
given the multivariate nature of their involvement.  For classification methods
such as Fisher's linear descriminate analysis (LDA),which maximally seperates
class means given their standard deviation, feature weights are representative
of the differences between class means. For support vector classification,
classifier weights are derived from observations that are the most similar
between classes, and hence represents the border between classes.  It is perhaps
most reasonable to adapt a conservative interpretation in which MVPA is used to
identify the most relevant features that are later tested in followon
experiments better suited to elucidating their relationship to the variable of
interest.

\emph{ --- CC text from TICS paper, do not use without modification!}
Multivariate Prediction Analysis. When applied to the study of intrinsic
activity, the goal of discovery science is to identify models that relate
measures of that activity (such as iFC) to phenotypic variables.  Prediction
analysis provides a means for measuring how well these models generalize to
independent data.  This is complementary to inferential statistics, which
measure the likelihood of such relationships arising by chance. In the
prediction analysis framework, a model relating iFC to a phenotype is learned
from a training dataset. This model is then applied to an independent test
dataset to predict phenotypes.  The resultant predictions are compared to the
true phenotypes to estimate how well the model generalizes to the tese dataset.
Thus, prediction analysis provides a natural framework for evaluating biomarkers
[21], performing real-time fMRI [88], and evaluating experimental tradeoffs
[89].

Prediction analysis has been applied to functional neuroimaging data since the early 90’s [90], and more
recently to IFC data [91]. Most, if not all, analysis methods can be applied in a predictive modeling
framework but the majority of methods that have been applied to iFC are multivariate classification and
regression methods (referred to as MVPA; multivariate prediction analysis). Multivariate methods are more
sensitive to distributed patterns of iFC than their univariate counterparts. Additionally they provide a
means for evaluating the significance of an entire pattern using a single statistic, obviating the need
to correct for multiple comparisons.

Although there are many circumstances in which high prediction accuracy is the ultimate goal of an
analysis (e.g. predicting treatment outcome), in general it is desirable that the model also be
interpretable. Identifying the iFC measures (features) that are most important to the model is
problematic and an open issue for MVPA research. Several feature selection algorithms have been proposed
to address this issue, but there is no consensus on which is best [21]. We note that feature selection
methods that rely on feature-by-feature statistical tests require correction for multiple comparisons.

MVPA classification has already been successfully used to identify potential iFC biomarkers of Alzheimer’s
disease [92], major depression [93], schizophrenia [94], and autism [47], among others. MVPA
classification and regression techniques have also been applied to identify biomarkers of age [95,96]
and recent work has shown the utility of MVPA methods for deriving iFC models at the individual
level [97]. An in-depth overview of the statistical pattern recognition methods underlying MVPA
techniques can be found in [98].
\emph{ --- END CC text from TICS paper, do not use without modification!}

>>>>

\subsection{Predictive Modeling}

\cite{craddock2009}
craddock 2009, Dosenbach 2010, lord 2011, deshpande,

@@

% Define predictive modeling and give examples

% Feature extraction and parameter selection

% Importance and caveats of interpretation

@@

Nearly every analytical method can be applied in a predictive modeling framework, but
multivariate classification and regression have been the most commonly used
(referred to as MVPA; multivariate prediction analysis).  Multivariate methods
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisions since
the significance of an entire pattern is evaluated using a single statistical
test. Predictive modeling is typically performed in a supervised learning
framework where a target is available for each observation of the features,
although unsupervised and semi-supervised methods can also be used.

@@

Model training can be further broken down into feature extraction, model
selection and feature selection and testing can be performed for model
optimization and

@@

Determining the subset of features most relevant to the model (feature
selection) is essential for interpreting the models, and is the subject of
ongoing research. In cases were model weights can be directly mapped to their
corresponding feature, there is typically no statistical framework for
thresholding these weights, and meaningless features are not always set to zero
(this is particularly the case for L2 regularized algorithms such as SVM and
ridge regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest using a
univariate statistical criterion such as t-test, Pearson's correlation, or
mutual information.  A drawback of this approach is that since the filters are
typically insensitive to multivaratiate information, they may exclude features
that would be important to the model.  Another approach is to treat feature
selection as a model selection problem in which case the features are
iteratively pruned and a subset is chosen based on prediction performance. A
third approach is to use sparse modeling methods such as L1 SVM or LASSO that
force meaningless weights to be zero.  Although ideal, these methods may have
limits on the number of non-zero weights used in the model or the exclusion of
highly correlated features from the model.  Contrary to previous reports [Guyon
2002, Craddock 2009] no matter which feature selection strategy is employed, it
must be performed with cross-validation to avoid biasing estimates of
prediction accuracy [later Guyon].

The interpretation of classifier weights for selected features, when available,
is complicated and requires insight into the mechanisms underlying the modeling
approach. If features are appropriately standardized prior to training, the
magnitude of weights can be interpreted to reflect the relative importance of
the feature to the model, but interpreting how the feature differs between
classes or relates to a continuous variable is more complicated, particularly
given the multivariate nature of their involvement.  For classification methods
such as Fisher's linear discriminant analysis (LDA), which maximally seperates
class means given their standard deviation, feature weights are representative
of the differences between class means. For support vector classification,
classifier weights are derived from observations that are the most similar
between classes, and hence represents the border between classes.  It is perhaps
most reasonable to adapt a conservative interpretation in which MVPA is used to
identify the most relevant features that are later tested in followon
experiments better suited to elucidating their relationship to the variable of
interest.
